Constraining gravitational wave amplitude birefringence with GWTC-3

### 2023-01-30 src/tex/ms.tex
collective results

The PE results are published in \citet{dataset}.

@@

We assume $\kappa$ is the same for all events and calculate a posterior distribution of $\kappa$.
The posterior distribution is given by
\begin{equation}
    p(\kappa|\{d\})\propto \prod_{i}p(\kappa|d_i)\,,
    \label{eq:restricted_posterior}
\end{equation}
where $p$ is the probability density function, $i$ is the index of the event, $\{d\}$ is the set of data including all events, and $d_i$ is the data of a single event.
Equation \ref{eq:restricted_posterior} shows that the posterior distribution of $\kappa$ is the product of the posterior distribution of $\kappa$ for each event.
This posterior distribution is a restricted posterior distribution because we assume $\kappa$ is the same for all events.

Note that we also perform a sanity test to check if our PE could recover the PE results done by LVK in \citet{GWTC-2.1, GWTC-3}.
To implement the sanity check, we set $\kappa$ to be $0$, which forces the PE results to be consistent with GR.
These sanity test results are also attached in \citet{dataset}.

% Bayesian Hierarchical Modeling
We then perform Bayesian hierarchical modeling to further analyze the constraint on $\kappa$.
Bayesian hierarchical modeling is a method to check if our waveform model can fully describe the amplitude birefringence effect.
Here, instead of assuming all events share the same value of $\kappa$, we assume $\kappa$ from each event is a true value drawn from a Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.
We chose a Gaussian distribution as the underlying distribution of $\kappa$ because it captures the correct mean and the standard deviation of the distribution no matter what the underlying distribution is.

>>>>

In order to validate our \ac{PE} implementation, we reproduce the \ac{LVK} \ac{PE} results obtained assuming \ac{GR} by enforcing $\kappa = 0$; this also has the advantage of producing \ac{GR} runs that are directly comparable to our birefringent runs.
All our \ac{PE} results, including the \ac{GR} valudation runs, are published in \citet{dataset}.

@@

\subsubsection{Shared birefringence parameter}

Since birefringence is a property of spacetime and we are targeting a propagation effect not intrinsic to any source, we should expect $\kappa$ to take the same value for all signals, whether it vanishes or not.
\mi{caveats?}
Under this assumption, the posterior on $\kappa$ inferred collectively from all events with a uniform prior is simply the product of the individual posteriors,%
\footnote{Had the $\kappa$ prior been nonuniform, then the individual \emph{likelihoods} would need to be multiplied, as opposed to the posteriors.}
\begin{equation}
    p(\kappa \mid \{d_i\})\propto \prod_{i}p(\kappa \mid d_i)\,,
    \label{eq:restricted_posterior}
\end{equation}
where $d_i$ is the strain data for the $i$\textsuperscript{th} event.
We use this expression to obtain the primary constraint presented in this work.

\subsubsection{Nonshared birefringence parameters}

% Bayesian Hierarchical Modeling
In order to further characterize our set of measurements, and identify potential outliers, we carry out an additional collective analysis that allows $\kappa$ to take different values for each events.
To do this, we make use of hierarchical Bayesian inference to model the distribuion of $\kappa$'s consistent with the observed data.
Here, instead of assuming all events share the same value of $\kappa$, we assume $\kappa$ from each event is a true value drawn from a Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.
We chose a Gaussian distribution as the underlying distribution of $\kappa$ because it captures the correct mean and the standard deviation of the distribution no matter what the underlying distribution is.
